---
title: "<span style='color:#3498db;'>EDA Report</span> <br> <small style='font-size:0.6em; color:#7f8c8d;'>Loan Approval Prediction</small>"
author: "<i class='fas fa-user'></i> **Metodi Latinov**"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cosmo
    highlight: tango
    css: styles.css
params:
  eda_data: NULL  # Required for targets data passing
knit: targets::tar_knit  # Must be at root level
---

```{r setup, include=FALSE}
loan_train_eda <- params$eda_data
```

# 1.Univariate Analysis


## Summary statistics 1.1


```{r}
datatable(summary_table_stat(data = loan_train_eda))
```


## 1.2 Boxplot

```{r}
boxplot_function(
  data = loan_train_eda,
  col_names = c("person_age","person_income","person_emp_length",
                "loan_amnt","loan_int_rate","loan_percent_income",
                "cb_person_cred_hist_length"))
```

## 1.3 Barplot 

```{r}
bar_plot_function(
  data = loan_train_eda,
  cols = c("person_home_ownership","loan_intent","loan_grade","cb_person_default_on_file",
           "loan_status"))
```



# 2.Bivariate Analysis

## Loan amount trends with income

### 2.1.1 Scatter plot Loan Amount X Personal Income

```{r message=TRUE, warning=FALSE}
scatter_plot(
  data =loan_train_eda,
  feature_x = "person_income",
  feature_y = "loan_amnt")
```

## Relantionship Between Age and Person Employment Lentht

### 2.1.2 Scatter plot Person Age X Person Employment Lenght

```{r message=TRUE, warning=FALSE}
scatter_plot(
  data =loan_train_eda,
  feature_x = "person_age",
  feature_y = "person_emp_length")
```


## Relantionship Between Age and Credict History Lenght

### 2.1.3 Scatter plot Person Age X Credict History Lenght

```{r message=TRUE, warning=FALSE}
scatter_plot(
  data =loan_train_eda,
  feature_x = "person_age",
  feature_y = "cb_person_cred_hist_length")
```


## 	Older applicants and interest rates?

### 2.1.4 Scatter plot Person Age X Loan Interest Rates
```{r message=TRUE, warning=FALSE}
scatter_plot(
  data =loan_train_eda,
  feature_x = "person_age",
  feature_y = "loan_int_rate")
```


## Default rate by credit grade

### 2.2.1 Stacked Bar Plot : Loan Grade X Loan Status

```{r}
stacked_bar_plot(
  data = loan_train_eda,
  feature = "loan_grade",
  fill = "loan_status")
```


## Impact of home ownership on default

### 2.2.2 Stacked Bar Plot : Home Ownership X Loan Status

```{r}
stacked_bar_plot(
  data = loan_train_eda,
  feature = "person_home_ownership",
  fill = "loan_status")
```

## Purpose of loan and repayment behavior

### 2.2.3 Stacked Bar Plot : Perpose of loan X Loan status

```{r}
stacked_bar_plot(
  data = loan_train_eda,
  feature = "loan_intent",
  fill = "loan_status")
```

## Prior default as a risk factor

### 2.2.4 Stacked Bar Plot : Has a record of defaulting on a loan X Loan Status

```{r}
stacked_bar_plot(
  data = loan_train_eda,
  feature = "cb_person_default_on_file",
  fill = "loan_status")
```

## Do defaulters have higher interest rates?

### 2.3.1 Bar Plot of means & Boxplot : Interest Rates X Loan Status

```{r}
bar_plot_means_boxplot(
  data = loan_train_eda,
  feature = "loan_int_rate",
  category = "loan_status" )
```


## Do defaulters borrow more?

### 2.3.2 Bar Plot of means & Boxplot : Interest Rates X Loan Status

```{r}
bar_plot_means_boxplot(
  data = loan_train_eda,
  feature = "loan_amnt",
  category = "loan_status")
```

## Employment stability and loan outcomes 

### 2.3.3 Bar Plot of means & Boxplot : Person Emplyment Lenght X Loan Status

```{r}
bar_plot_means_boxplot(
  data = loan_train_eda,
  feature = "person_emp_length",
  category = "loan_status")
```

## Interest rates by grade

### 2.3.4 Bar Plot of means & Boxplot : Loan Grade X Intrest Rates

```{r}
bar_plot_means_boxplot(
  data = loan_train_eda,
  feature = "loan_int_rate",
  category = "loan_grade")
```

## Loan sizes by grade

### 2.3.5 Bar Plot of means & Boxplot : Loan Grade X Loan Amount

```{r}
bar_plot_means_boxplot(
  data = loan_train_eda,
  feature = "loan_amnt",
  category = "loan_grade")
```

## Default rate by grade and intent

### 2.4.1 Heatmap : Loan Grade X Loan Intent X Loan Status

```{r}
loan_train_eda %>%
  group_by(loan_grade, loan_intent) %>%
  summarise(count_status = n(), .groups = "drop")%>%
  heatmap_function(
    feature_amount = "count_status",
    category_x = "loan_grade",
    category_y = "loan_intent")
```


